# Truthful, Visually-Driven Survey Analysis Income of Rural Informal Labourers in Northern Vietnam

## 1. Objective and dataset 

This report delivers 5–10 visually-driven observations from a rural informal labour survey in Northern Vietnam, using static ggplot2 visualizations . The dataset contains 725 rows and 30 variables, and a preliminary audit detects 2 duplicated rows. A key measurement property is validated: wherever all components are present, TEIN = TAIN + TSII + TOIN holds exactly . This su›pports truthful “income composition” visuals.

### 1.1 Data Preparation

```{r}
#| message: false
#| warning: false
pacman::p_load(
  tidyverse, janitor, skimr,
  scales, patchwork,
  ggridges, ggdist,
  sf, rnaturalearth, rnaturalearthdata,
  ggrepel
)

set.seed(1234)

# Required variable name per prompt
elsiver <- read_csv("data/Upload for elsiver.csv", show_col_types = FALSE)

dim(elsiver)
glimpse(elsiver)

```

Two fully duplicated records are detected. This is small but should be removed to avoid double-counting in subgroup summaries.

```{r}
#| message: false
#| warning: false
# Duplicate rows
dup_n <- sum(duplicated(elsiver))
dup_n

# Missingness (counts and %)
miss_tbl <- elsiver %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "missing_n") %>%
  mutate(
    n = nrow(elsiver),
    missing_pct = missing_n / n,
    domain = case_when(
      str_starts(var, "C") ~ "C (Characteristics)",
      str_starts(var, "T") ~ "T (Income components)",
      str_starts(var, "F") ~ "F (Income factors)",
      str_starts(var, "L") ~ "L (Living conditions)",
      str_starts(var, "P") ~ "P (Policies)",
      str_starts(var, "A") ~ "A (Assessments)",
      TRUE ~ "Other"
    )
  ) %>%
  arrange(desc(missing_pct))

miss_tbl %>% slice_head(n = 12)

```

```{r}
#| message: false
#| warning: false
# TEIN consistency check: TEIN == TAIN + TSII + TOIN wherever all are present
tein_check <- elsiver %>%
  filter(!is.na(TEIN), !is.na(TAIN), !is.na(TSII), !is.na(TOIN)) %>%
  mutate(diff = TEIN - (TAIN + TSII + TOIN))

tein_check %>%
  summarise(
    n_checked = n(),
    max_abs_diff = max(abs(diff)),
    all_exact_zero = all(diff == 0)
  )

```

```{r}
#| message: false
#| warning: false
elsiver <- elsiver %>%
  mutate(
    # Fix known parsing issues / out-of-range anomalies
    ARO4 = if_else(ARO4 == "`", NA_character_, as.character(ARO4)),
    ARO4 = readr::parse_number(ARO4),
    ARO2 = as.numeric(ARO2),
    ARO2 = if_else(ARO2 %in% c(1, 2, 3), ARO2, NA_real_),  # assumption: valid range is 1–3 here

    PPO1 = if_else(PPO1 == 0, NA_real_, PPO1),
    PPO2 = if_else(PPO2 == 0, NA_real_, PPO2),

    # Impute TSII missing as 0 only when TEIN == TAIN + TOIN (arithmetically implied)
    TSII = case_when(
      is.na(TSII) &
        !is.na(TEIN) & !is.na(TAIN) & !is.na(TOIN) &
        abs(TEIN - (TAIN + TOIN)) < 1e-9 ~ 0,
      TRUE ~ TSII
    ),

    # Derived, analysis-friendly variables
    CPRO = factor(CPRO, levels = 1:5, labels = paste("Province code", 1:5)),

    CGEN = factor(CGEN, levels = c(1, 2), labels = c("Male", "Female")),
    CRAC = factor(CRAC, levels = c(1, 2), labels = c("Kinh", "Ethnic minorities")),
    CJOB = factor(CJOB, levels = c(1, 2, 3), labels = c("Agriculture", "Service/Industrial", "Other")),
    CJOB = fct_explicit_na(CJOB, "Missing (CJOB)"),

    CQUI = factor(CQUI, levels = c(1, 2, 3, 4, 5),
                  labels = c("Top quintile", "Fourth quintile", "Middle quintile", "Second quintile", "Lowest quintile")),
    CQUI = fct_explicit_na(CQUI, "Missing (CQUI)"),

    FEDU = factor(FEDU, levels = c(0, 1, 2, 3),
                  labels = c("Primary", "Lower secondary", "Upper secondary", "Other / higher")),
    FVTP = factor(FVTP, levels = c(1, 2), labels = c("Short courses", "Long courses")),
    FVTP = fct_explicit_na(FVTP, "Missing (FVTP)"),

    FCRA = factor(FCRA, levels = c(1, 2), labels = c("Credit inaccessible", "Credit accessible")),
    FCRA = fct_explicit_na(FCRA, "Missing (FCRA)"),

    FTAP = factor(FTAP, levels = c(1, 2), labels = c("Low technique", "Adequate technique")),
    FTAP = fct_explicit_na(FTAP, "Missing (FTAP)"),

    # Housing type derived from one-hot indicators
    house_type = case_when(
      LHO1 == 1 ~ "Cottage house",
      LHO2 == 1 ~ "Roofed house",
      LHO3 == 1 ~ "Solid house",
      LHO4 == 1 ~ "Building",
      TRUE ~ "Missing/Unknown"
    ) %>% factor(levels = c("Cottage house", "Roofed house", "Solid house", "Building", "Missing/Unknown")),

    # log1p for skewed income (keeps zeros defined)
    log1p_TEIN = log1p(TEIN)
  )

```

Missingness is heavily concentrated in housing/loan/savings fields (e.g., LCRE, LSAV, LHO1–LHO4). Importantly, housing indicators behave like a **one-choice encoding**: most respondents have **exactly one** of LHO1–LHO4 filled, while others remain NA. This is **structural missingness**, not “non-response”.Treating these NAs as “missing at random” and applying `drop_na()` would bias the sample and distort subgroup comparisons.

```{r}
#| fig-width: 10
#| fig-height: 6
miss_plot_df <- elsiver %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "missing_n") %>%
  mutate(
    n = nrow(elsiver),
    missing_pct = missing_n / n,
    domain = case_when(
      str_starts(var, "C") ~ "C",
      str_starts(var, "T") ~ "T",
      str_starts(var, "F") ~ "F",
      str_starts(var, "L") ~ "L",
      str_starts(var, "P") ~ "P",
      str_starts(var, "A") ~ "A",
      TRUE ~ "Other"
    ),
    var = fct_reorder(var, missing_pct)
  )

ggplot(miss_plot_df, aes(x = var, y = missing_pct)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  facet_wrap(~ domain, scales = "free_y") +
  labs(
    title = "Missingness by variable (faceted by survey domain)",
    x = NULL,
    y = "Missing (% of 725 rows)"
  )

```

### Figure 2 — Total income (TEIN) distribution (P99 zoom + log1p view)

TEIN is heavily right-skewed: most observations cluster at low-to-moderate income, while a small number of high-income cases generate a long tail. This makes “average income” an unstable summary; median and distribution-aware plots are more truthful than a single mean. The paired panels follow best-practice guidance: the left panel is a zoomed view to show the bulk distribution ; the right panel uses log1p to show tail structure

```{r}
#| fig-width: 10
#| fig-height: 4.5
tein_p99 <- quantile(elsiver$TEIN, 0.99, na.rm = TRUE)
tein_med <- median(elsiver$TEIN, na.rm = TRUE)

p_lin <- elsiver %>%
  filter(!is.na(TEIN)) %>%
  ggplot(aes(x = TEIN)) +
  geom_histogram(binwidth = 5, boundary = 0) +
  geom_vline(xintercept = tein_med, linetype = "dashed") +
  coord_cartesian(xlim = c(0, tein_p99)) +
  labs(
    title = "Total income (TEIN) distribution (zoomed to P99)",
    subtitle = paste0("Dashed line = median; unit = Mil VND; zoom avoids deceptive truncation"),
    x = "TEIN (Mil VND)",
    y = "Count"
  )

p_log <- elsiver %>%
  filter(!is.na(TEIN)) %>%
  ggplot(aes(x = TEIN)) +
  geom_histogram(binwidth = 0.25) +
  scale_x_continuous(trans = "log1p") +
  labs(
    title = "Same data on log1p scale",
    subtitle = "log1p handles TEIN=0 and makes the long tail visible",
    x = "TEIN (Mil VND), log1p scale",
    y = "Count"
  )

p_lin + p_log

```

\

### Income by education using a raincloud design

```{r}
library(dplyr)
library(forcats)
library(ggplot2)
library(ggdist)

# CORRECTED raincloud plot code
df <- elsiver %>%
  filter(!is.na(TEIN), !is.na(FEDU)) 

p99 <- quantile(df$TEIN, 0.99, na.rm = TRUE)

ggplot(df, aes(y = FEDU, x = TEIN, fill = FEDU)) +
  stat_halfeye(
    adjust = 0.8,
    justification = -0.25,
    .width = c(0.5, 0.8, 0.95),
    point_interval = "median_qi",
    alpha = 0.6
  ) +
  geom_boxplot(width = 0.18, outlier.alpha = 0.25, alpha = 0.3) +
  geom_point(
    position = position_jitter(height = 0.08, width = 0),
    alpha = 0.08,
    size = 0.8
  ) +
  coord_cartesian(xlim = c(0, p99)) +
  labs(
    title = "Income differs strongly by education (distribution-first view)",
    subtitle = "Half-eye + box + jitter (raincloud) shows distribution and uncertainty",
    x = "TEIN (Mil VND), zoomed to P99",
    y = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

Income rises with education in both central tendency and upper-tail reach: higher education groups show higher medians and progressively wider distributions. Crucially, the distribution shapes matter: the most educated group is small and has a dramatically longer tail. Higher education does not guarantee strictly higher *minimum* income, but is associated with much higher *maximum* income potential. 

Education is not only associated with higher income; it may also act as a gateway to **non-farm income sources**, which can be tested using the TEIN decomposition.

### Income by career with ridgelines, stratified by race

```{r}
#| out-width: "100%"
#| fig-retina: 2
library(dplyr)
library(ggplot2)
library(ggridges)

plot_df <- elsiver %>%
  mutate(
    log1p_TEIN = log1p(TEIN),
    CJOB = as.factor(CJOB),
    CRAC = as.factor(CRAC)
  ) %>%
  filter(!is.na(log1p_TEIN), !is.na(CJOB), !is.na(CRAC))

ggplot(plot_df, aes(x = log1p_TEIN)) +
  ggridges::geom_density_ridges(aes(y = CJOB, fill = CJOB),
                                alpha = 0.6, color = "white", scale = 1.1) +
  ggplot2::geom_boxplot(aes(y = CJOB),
                        width = 0.12, alpha = 0.25, outlier.alpha = 0.15) +
  facet_wrap(~ CRAC) +
  labs(
    title = "Career income distributions differ, and race stratification matters",
    subtitle = "Ridgelines + box overlays on log1p scale reveal distribution shifts and within-group spread",
    x = "log1p(TEIN) (Mil VND)",
    y = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

Across both race strata, agriculture has the lowest income distribution, while service/industrial and other careers shift upward. However, stratifying by race reveals a crucial inequality pattern: in agriculture, the ethnic-minority distribution is noticeably left-shifted relative to Kinh, indicating lower typical income among minority agricultural workers. This intersectional view is more truthful than a single “race gap” or “career gap” because it shows *where* disparities concentrate. \
Limitations: Density estimation can smooth away micro-structure, and snowball sampling may cluster respondents within ethnic and occupational networks. Therefore, the plot should be framed as exploratory evidence, not a population estimate.

### Median income composition by education using a waterfall design

```{r}
#| fig-width: 11
#| fig-height: 6
comp_med <- elsiver %>%
  filter(!is.na(TEIN), FEDU != "Missing (FEDU)") %>%
  group_by(FEDU) %>%
  summarise(
    TAIN_med = median(TAIN, na.rm = TRUE),
    TSII_med = median(TSII, na.rm = TRUE),
    TOIN_med = median(TOIN, na.rm = TRUE),
    TEIN_med = median(TEIN, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(TAIN_med, TSII_med, TOIN_med),
               names_to = "component", values_to = "value") %>%
  mutate(
    component = recode(component,
      TAIN_med = "Agricultural (TAIN)",
      TSII_med = "Service/Industrial (TSII)",
      TOIN_med = "Other income (TOIN)"
    ),
    component = factor(component, levels = c("Agricultural (TAIN)", "Service/Industrial (TSII)", "Other income (TOIN)"))
  ) %>%
  group_by(FEDU) %>%
  mutate(
    start = lag(cumsum(value), default = 0),
    end = start + value
  ) %>%
  ungroup()

ggplot(comp_med, aes(y = component, x = value)) +
  geom_col(aes(fill = component), width = 0.7) +
  geom_segment(aes(x = start, xend = end, y = component, yend = component), linewidth = 1) +
  facet_wrap(~ FEDU, scales = "free_x") +
  labs(
    title = "How income composition shifts with education (median-based waterfall)",
    subtitle = "Waterfall components sum to median TEIN because TEIN identity holds in the data",
    x = "Median component income (Mil VND)",
    y = NULL
  ) +
  theme(legend.position = "none")

```

Education is not only associated with higher total income; it is associated with a different *composition* of income. At primary and lower-secondary levels, median income is dominated by agricultural income (TAIN), with minimal service/industrial contribution. As education increases, non-agricultural components (TSII and TOIN) become much larger in the median composition, consistent with a broader opportunity set beyond agriculture. Because the dataset satisfies TEIN = TAIN + TSII + TOIN exactly (where complete), a waterfall is a truthful decomposition rather than an illustrative approximation. \
Limitations: This is a median-based decomposition (not a population budget). The highest education group has small N, and snowball sampling limits external generalization. 

### Race × career interaction as an annotated heatmap

```{r}
#| fig-width: 9
#| fig-height: 4.5
heat_df <- elsiver %>%
  filter(!is.na(TEIN), CJOB != "Missing (CJOB)") %>%
  group_by(CRAC, CJOB) %>%
  summarise(
    n = n(),
    tein_med = median(TEIN),
    .groups = "drop"
  )

ggplot(heat_df, aes(x = CJOB, y = CRAC, fill = tein_med)) +
  geom_tile() +
  geom_text(aes(label = paste0("Median=", round(tein_med, 1), "\nN=", n)), size = 3) +
  labs(
    title = "Intersectional view: race-by-career median income (with sample size)",
    x = NULL,
    y = NULL,
    fill = "Median TEIN\n(Mil VND)"
  )

```

This heatmap compresses an intersectional story into a transparent summary by pairing medians with sample sizes. The largest disparity appears in agriculture: Kinh agricultural respondents have a substantially higher median income than ethnic-minority agricultural respondents, while non-agricultural careers show smaller differences (and smaller N). Reporting both “Median” and “N” reduces the risk of overinterpreting thin cells, and helps keep the conclusions proportional to evidence. \
Limitations: Medians suppress within-cell distribution spread; unequal cell sizes and snowball sampling imply that these are sample summaries, not population estimates, and cannot establish causal discrimination mechanisms. 

### Loan amount vs income among credit-access respondents

```{r}
#| fig-width: 9
#| fig-height: 4.5
elsiver %>%
  
  ggplot(aes(x = LCRE, y = TEIN)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_y_continuous(trans = "log1p") +
  labs(
    title = "Among credit-access respondents, larger loans align with higher income",
    subtitle = "Conditional subgroup view (FCRA = accessible); TEIN shown on log1p scale",
    x = "Loan amount (LCRE, Mil VND)",
    y = "TEIN (Mil VND), log1p scale"
  )

```

Within the credit-access subgroup (where loan amount is observed by design), higher reported loans are associated with higher reported income. This may reflect productive investment (e.g., upgrading tools/technology) or selection effects (higher-income individuals are more likely to qualify for larger loans), both of which are consistent with the paper’s framing that credit supports productivity improvements but is also constrained by rural savings limitations. \
Limitations: This is a conditional analysis on a small subgroup and is not causal. Credit access is not randomized, and snowball sampling plus eligibility constraints can create strong selection bias. 

The survey covers five provinces in northern Vietnam’s mountainous region (highlighted). A context map supports truthfulness by grounding the analysis geographically for readers unfamiliar with the region and reducing the risk of interpreting the dataset as nationally representative. Because the dataset documentation does not specify how numeric `CPRO`codes map to these province names, this figure intentionally avoids joining survey records to province polygons. \
Limitations: This is a contextual visualization. Any province-level income comparisons by code should be labelled as “province code” unless an official mapping is obtained.

```{r}
#| fig-width: 9
#| fig-height: 6

library(maps)
library(ggplot2)
library(dplyr)

# Vietnam boundary (data.frame)
vietnam_map <- map_data("world") %>%
  filter(region == "Vietnam")

# Northern Vietnam bounding box (data.frame, not sf)
north_box_df <- data.frame(
  long = c(102, 108, 108, 102, 102),
  lat  = c(20,  20,  24,  24,  20)
)

ggplot() +
  geom_polygon(
    data = vietnam_map,
    aes(x = long, y = lat, group = group),
    fill = "grey90",
    color = "grey50",
    linewidth = 0.3
  ) +
  geom_polygon(
    data = north_box_df,
    aes(x = long, y = lat),
    fill = "#2b8cbe",
    alpha = 0.25,
    color = "#2b8cbe",
    linewidth = 0.6
  ) +
  coord_fixed(1.3) +
  labs(
    title = "Survey geography context: Northern Vietnam highlighted",
    subtitle = "Context map only; does not join to CPRO codes (code→name mapping not verified)",
    caption = "Source: maps package"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold")
  )
```
