# Truthful, Visually-Driven Survey Analysis Income of Rural Informal Labourers in Northern Vietnam

## 1. Objective and dataset 

This report delivers 5–10 visually-driven observations from a rural informal labour survey in Northern Vietnam, using static ggplot2 visualizations . The dataset contains 725 rows and 30 variables, and a preliminary audit detects 2 duplicated rows. A key measurement property is validated: wherever all components are present, TEIN = TAIN + TSII + TOIN holds exactly . This su›pports truthful “income composition” visuals.

### 1.1 Data Preparation

```{r}
#| message: false
#| warning: false
pacman::p_load(
  tidyverse, janitor, skimr,
  scales, patchwork,
  ggridges, ggdist,
  sf, rnaturalearth, rnaturalearthdata,
  ggrepel
)

set.seed(1234)
elsiver <- read_csv("data/Upload for elsiver.csv", show_col_types = FALSE)

dim(elsiver)
glimpse(elsiver)

```

Two fully duplicated records are detected. This is small but should be removed to avoid double-counting in subgroup summaries.

```{r}
#| message: false
#| warning: false
dup_n <- sum(duplicated(elsiver))
dup_n

miss_tbl <- elsiver %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "missing_n") %>%
  mutate(
    n = nrow(elsiver),
    missing_pct = missing_n / n,
    domain = case_when(
      str_starts(var, "C") ~ "C (Characteristics)",
      str_starts(var, "T") ~ "T (Income components)",
      str_starts(var, "F") ~ "F (Income factors)",
      str_starts(var, "L") ~ "L (Living conditions)",
      str_starts(var, "P") ~ "P (Policies)",
      str_starts(var, "A") ~ "A (Assessments)",
      TRUE ~ "Other"
    )
  ) %>%
  arrange(desc(missing_pct))

miss_tbl %>% slice_head(n = 12)

```

```{r}
#| message: false
#| warning: false
# TEIN == TAIN + TSII + TOIN 
tein_check <- elsiver %>%
  filter(!is.na(TEIN), !is.na(TAIN), !is.na(TSII), !is.na(TOIN)) %>%
  mutate(diff = TEIN - (TAIN + TSII + TOIN))

tein_check %>%
  summarise(
    n_checked = n(),
    max_abs_diff = max(abs(diff)),
    all_exact_zero = all(diff == 0)
  )

```

```{r}
#| message: false
#| warning: false
elsiver <- elsiver %>%
  mutate(
    ARO4 = if_else(ARO4 == "`",NA_character_, as.character(ARO4)),
    ARO4 = readr::parse_number(ARO4),
    ARO2 = as.numeric(ARO2),
    ARO2 = if_else(ARO2 %in% c(1,2,3), ARO2, NA_real_),  # assumption: valid range is 1–3 here

    PPO1 = if_else(PPO1 == 0,NA_real_, PPO1),
    PPO2 = if_else(PPO2 == 0, NA_real_, PPO2),


    TSII = case_when(
      is.na(TSII) &
        !is.na(TEIN) & !is.na(TAIN) & !is.na(TOIN) &
        abs(TEIN - (TAIN + TOIN)) < 1e-9 ~ 0,
      TRUE ~ TSII
    ),

    
    CPRO = factor(CPRO, levels = 1:5, labels = paste("Province code", 1:5)),

    CGEN = factor(CGEN, levels = c(1, 2), labels = c("Male", "Female")),
    CRAC = factor(CRAC, levels = c(1, 2), labels = c("Kinh", "Ethnic minorities")),
    CJOB = factor(CJOB, levels = c(1, 2, 3), labels = c("Agriculture", "Service/Industrial", "Other")),
    CJOB = fct_explicit_na(CJOB, "Missing (CJOB)"),

    CQUI = factor(CQUI, levels = c(1, 2, 3, 4, 5),
                  labels = c("Top quintile", "Fourth quintile", "Middle quintile", "Second quintile", "Lowest quintile")),
    CQUI = fct_explicit_na(CQUI, "Missing (CQUI)"),

    FEDU = factor(FEDU, levels = c(0, 1, 2, 3),
                  labels = c("Primary", "Lower secondary", "Upper secondary", "Other / higher")),
    FVTP = factor(FVTP, levels = c(1, 2), labels = c("Short courses", "Long courses")),
    FVTP = fct_explicit_na(FVTP, "Missing (FVTP)"),

    FCRA = factor(FCRA, levels = c(1, 2), labels = c("Credit inaccessible", "Credit accessible")),
    FCRA = fct_explicit_na(FCRA, "Missing (FCRA)"),

    FTAP = factor(FTAP, levels = c(1, 2), labels = c("Low technique", "Adequate technique")),
    FTAP = fct_explicit_na(FTAP, "Missing (FTAP)"),

    house_type = case_when(
      LHO1 == 1 ~ "Cottage house",
      LHO2 == 1 ~ "Roofed house",
      LHO3 == 1 ~ "Solid house",
      LHO4 == 1 ~ "Building",
      TRUE ~ "Missing/Unknown"
    ) %>% factor(levels = c("Cottage house", "Roofed house", "Solid house", "Building", "Missing/Unknown")),

    # log1p for skewed income
    log1p_TEIN = log1p(TEIN)
  )

```

Missingness is heavily concentrated in living-condition variables (LCRE, LHO1–LHO4, LSAV), with some exceeding 70–90% missingness.However, housing variables behave like one-choice encoding (one of four filled), indicating structural missingness, not random non-response.

Additionally, income identity holds exactly:

TEIN = TAIN + TSII + TOIN (validated on 719 complete cases)

```{r}
#| fig-width: 10
#| fig-height: 6
miss_plot_df <- elsiver %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "missing_n") %>%
  mutate(
    n = nrow(elsiver),
    missing_pct = missing_n / n,
    domain = case_when(
      str_starts(var, "C") ~ "C",
      str_starts(var, "T") ~ "T",
      str_starts(var, "F") ~ "F",
      str_starts(var, "L") ~ "L",
      str_starts(var, "P") ~ "P",
      str_starts(var, "A") ~ "A",
      TRUE ~ "Other"
    ),
    var = fct_reorder(var, missing_pct)
  )

ggplot(miss_plot_df, aes(x = var, y = missing_pct)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  facet_wrap(~ domain, scales = "free_y") +
  labs(
    title = "Missingness by variable (faceted by survey domain)",
    x = NULL,
    y = "Missing (% of 725 rows)"
  )

```

## 2.Data Analyse

#### Figure 1 — Total income (TEIN) distribution (P99 zoom + log1p view)

TEIN is heavily right-skewed: most observations cluster at low-to-moderate income, while a small number of high-income cases generate a long upper tail. Mean income would exaggerate typical earnings. The paired panels follow best-practice guidance: the left panel is a zoomed view to show the bulk distribution ; the right panel uses log1p to show tail structure

```{r}
#| fig-width: 10
#| fig-height: 4.5
tein_p99 <- quantile(elsiver$TEIN, 0.99, na.rm = TRUE)
tein_med <- median(elsiver$TEIN, na.rm = TRUE)

p_lin <- elsiver %>%
  filter(!is.na(TEIN)) %>%
  ggplot(aes(x = TEIN)) +
  geom_histogram(binwidth = 5, boundary = 0) +
  geom_vline(xintercept = tein_med, linetype = "dashed") +
  coord_cartesian(xlim = c(0, tein_p99)) +
  labs(
    title = "Total income (TEIN) distribution (zoomed to P99)",
    subtitle = paste0("Dashed line = median; unit = Mil VND; zoom avoids deceptive truncation"),
    x = "TEIN (Mil VND)",
    y = "Count"
  )

p_log <- elsiver %>%
  filter(!is.na(TEIN)) %>%
  ggplot(aes(x = TEIN)) +
  geom_histogram(binwidth = 0.25) +
  scale_x_continuous(trans = "log1p") +
  labs(
    title = "Same data on log1p scale",
    subtitle = "log1p handles TEIN=0 and makes the long tail visible",
    x = "TEIN (Mil VND), log1p scale",
    y = "Count"
  )

p_lin + p_log

```

###  2.1 Education as Primary Income Gradient

#### Figure 2: Raincloud by FEDU

```{r}
library(dplyr)
library(forcats)
library(ggplot2)
library(ggdist)


df <- elsiver %>%
  filter(!is.na(TEIN), !is.na(FEDU)) 

p99 <- quantile(df$TEIN, 0.99, na.rm = TRUE)

ggplot(df, aes(y = FEDU, x = TEIN, fill = FEDU)) +
  stat_halfeye(
    adjust = 0.8,
    justification = -0.25,
    .width = c(0.5, 0.8, 0.95),
    point_interval = "median_qi",
    alpha = 0.6
  ) +
  geom_boxplot(width = 0.18, outlier.alpha = 0.25, alpha = 0.3) +
  geom_point(
    position = position_jitter(height = 0.08, width = 0),
    alpha = 0.08,
    size = 0.8
  ) +
  coord_cartesian(xlim = c(0, p99)) +
  labs(
    title = "Income differs strongly by education (distribution-first view)",
    subtitle = "Half-eye + box + jitter (raincloud) shows distribution and uncertainty",
    x = "TEIN (Mil VND), zoomed to P99",
    y = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

Income rises with education in both central tendency and upper-tail reach: higher education groups show higher medians and progressively wider distributions. Importantly, higher education does not guarantee higher minimum income, but increases maximum potential.

Education is may not only associated with higher income; it may also act as a gateway to non-farm income sources, which can be tested using the TEIN decomposition.

### 2.2 Occupational Structure and Race Stratification

#### Figure 3: Ridgeline by CJOB × CRAC

```{r}
#| out-width: "100%"
#| fig-retina: 2
library(dplyr)
library(ggplot2)
library(ggridges)

plot_df <- elsiver %>%
  mutate(
    log1p_TEIN = log1p(TEIN),
    CJOB = as.factor(CJOB),
    CRAC = as.factor(CRAC)
  ) %>%
  filter(!is.na(log1p_TEIN), !is.na(CJOB), !is.na(CRAC))

ggplot(plot_df, aes(x = log1p_TEIN)) +
  ggridges::geom_density_ridges(aes(y = CJOB, fill = CJOB),
                                alpha = 0.6, color = "white", scale = 1.1) +
  ggplot2::geom_boxplot(aes(y = CJOB),
                        width = 0.12, alpha = 0.25, outlier.alpha = 0.15) +
  facet_wrap(~ CRAC) +
  labs(
    title = "Career income distributions differ, and race stratification matters",
    subtitle = "Ridgelines + box overlays on log1p scale reveal distribution shifts and within-group spread",
    x = "log1p(TEIN) (Mil VND)",
    y = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

Across both race strata, agriculture has the lowest income distribution, while service/industrial and other careers shift upward. However, Grouped by race reveals a crucial inequality pattern: in agriculture, the ethnic-minority distribution is noticeably left-shifted relative to Kinh, indicating lower typical income among minority agricultural workers.

### 2.3 Mechanism: Income Composition Shift

#### Figure 4: Waterfall by education

```{r}
#| fig-width: 11
#| fig-height: 6
comp_med <- elsiver %>%
  filter(!is.na(TEIN), FEDU != "Missing (FEDU)") %>%
  group_by(FEDU) %>%
  summarise(
    TAIN_med = median(TAIN, na.rm = TRUE),
    TSII_med = median(TSII, na.rm = TRUE),
    TOIN_med = median(TOIN, na.rm = TRUE),
    TEIN_med = median(TEIN, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(TAIN_med, TSII_med, TOIN_med),
               names_to = "component", values_to = "value") %>%
  mutate(
    component = recode(component,
      TAIN_med = "Agricultural (TAIN)",
      TSII_med = "Service/Industrial (TSII)",
      TOIN_med = "Other income (TOIN)"
    ),
    component = factor(component, levels = c("Agricultural (TAIN)", "Service/Industrial (TSII)", "Other income (TOIN)"))
  ) %>%
  group_by(FEDU) %>%
  mutate(
    start = lag(cumsum(value), default = 0),
    end = start + value
  ) %>%
  ungroup()

ggplot(comp_med, aes(y = component, x = value)) +
  geom_col(aes(fill = component), width = 0.7) +
  geom_segment(aes(x = start, xend = end, y = component, yend = component), linewidth = 1) +
  facet_wrap(~ FEDU, scales = "free_x") +
  labs(
    title = "How income composition shifts with education (median-based waterfall)",
    subtitle = "Waterfall components sum to median TEIN because TEIN identity holds in the data",
    x = "Median component income (Mil VND)",
    y = NULL
  ) +
  theme(legend.position = "none")

```

Education does not only increase total income — it changes income structure :

-   Primary/Lower secondary: median income dominated by agricultural income.

-   Higher education: non-farm components (TSII, TOIN) increase substantially.

This provides a mechanism:

Education → Access to non-farm activities → Diversified income → Higher TEIN

### 2.4 Intersectional Inequality

#### Figure 5: Race × Career Heatmap

```{r}
#| fig-width: 9
#| fig-height: 4.5
heat_df <- elsiver %>%
  filter(!is.na(TEIN), CJOB != "Missing (CJOB)") %>%
  group_by(CRAC, CJOB) %>%
  summarise(
    n = n(),
    tein_med = median(TEIN),
    .groups = "drop"
  )

ggplot(heat_df, aes(x = CJOB, y = CRAC, fill = tein_med)) +
  geom_tile() +
  geom_text(aes(label = paste0("Median=", round(tein_med, 1), "\nN=", n)), size = 3) +
  labs(
    title = "Intersectional view: race-by-career median income (with sample size)",
    x = NULL,
    y = NULL,
    fill = "Median TEIN\n(Mil VND)"
  )

```

Largest disparity occurs in agriculture. Kinh agricultural respondents have significantly higher median income than ethnic minorities. Non-agricultural cells show smaller differences but also smaller N. Race gap is structurally embedded in agricultural sector.

### 2.5 Credit and Income 

### Figure 6: Loan amount vs income among credit-access respondents

```{r}
#| fig-width: 9
#| fig-height: 4.5
elsiver %>%
  
  ggplot(aes(x = LCRE, y = TEIN)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_y_continuous(trans = "log1p") +
  labs(
    title = "Among credit-access respondents, larger loans align with higher income",
    subtitle = "Conditional subgroup view (FCRA = accessible); TEIN shown on log1p scale",
    x = "Loan amount (LCRE, Mil VND)",
    y = "TEIN (Mil VND), log1p scale"
  )

```

Within the credit-access subgroup (where loan amount is observed by design), higher reported loans are associated with higher reported income. This may reflect productive investment (e.g., upgrading tools/technology) or selection effects (higher-income individuals are more likely to qualify for larger loans), both of which are consistent with the paper’s framing that credit supports productivity improvements but is also constrained by rural savings limitations.

## 3.Conclusion

This analysis demonstrates that income inequality among rural informal labourers is structurally patterned rather than random.

First, income is highly right-skewed, making median-based and distribution-first visualisation essential for truthful interpretation. Second, Higher education is associated not only with higher total income but also with a shift from agricultural to non-farm income sources.Third, occupation plays a critical role. Agriculture consistently exhibits the lowest income distribution. Race disparities are concentrated within the agricultural sector. Finally, credit access is positively associated with income.

Overall, the findings suggest that policies aimed at improving educational attainment and facilitating transitions into non-agricultural activities may have the greatest potential to reduce income inequality in rural northern Vietnam.
