---
title: "Hands-on_Exercise2"
---

# Overview

**Summary:** This section introduces the main goals of the chapter. It tells us we will learn extensions for `ggplot2`. We will learn how to fix overlapping labels, use professional themes, and combine many plots into one figure.

## 1 Preparation

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes,
               tidyverse) 

```

```{r}
remotes::install_git("https://codeberg.org/hrbrmstr/hrbrthemes.git")
```

```{r setup, include=FALSE}
library(tidyverse)   
library(hrbrthemes)  
library(ggrepel)     
library(patchwork)   
```

```{r}
exam_data <- read_csv("Hands-on-Exercise2/data/Exam_data.csv")
```

## **2.3 ggrepel**

**This section solves a big problem: overlapping text. When we use geom_text, labels cover each other. The package ggrepel fixes this automatically.I can change the box.padding argument. This controls the space around the labels. I can also change the color of the labels to match the data points, making it look more organized.**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## 2.4.1 ggtheme package

**This section introduces built-in themes like `theme_gray()` and external themes like `theme_economist()` and `theme_ipsum()`. These themes change the background, fonts, and grid lines.**

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="powderblue") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

## 2.4.2 hrbthemes

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="lightblue") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

## 2.5 Beyond Single Graph

**Sometimes one chart is not enough. This section teaches patchwork. We can use +, /, and () to arrange multiple charts (side-by-side or stacked)**

graph1: single graph of math grade

```{r}
p1 <-ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
```

graph1: single graph of english grade

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
```

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

```{r}
(p3|(p1 / p2) ) + 
  plot_annotation(tag_levels = 'I')
```

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
